Visual Interrogation of Attention-Based Models for Natural Language Inference and Machine Comprehension

### 2018-06-01 EMNLP-demo-paper/result.tex
minor edits all over

\subsection{Natural Language Inference}
\label{sec:NLIexample}
The NLI task predicts the relationship between a premise sentence (P) and a hypothesis sentence (H).
The attention matrix captures the alignment of words between these two sentences.
Here we give an example of how a wrong prediction can be correct by editing the attention value.

@@

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pair (P:``A woman in a green jacket is drinking tea.'' H:``A woman is drinking green tea.'') is predicted to be \emph{entailment}, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two \emph{greens} decorate different nouns, which potentially leads to the wrong prediction. The grammar structure is visually presented by dependency tree.
%i.e., \emph{green} in \textbf{P} is attached to ``jacket''.
However, the model does not has access to the grammar information and mistakenly assumes the two \emph{greens} are describing the same thing.
% (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "greens" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

\subsection{Machine Comprehension}
\label{sec:MCexample}
In the machine comprehension task, the goal is to select a span of text as the answer
to a question sentence.
%two sequences of texts are given: context and question.
%The goal is to select a span of text from the context that answers the question. 
%We provide a simple model formulation and refer the reader to the paper for details.
%%\begin{align}
%	$C^\prime, Q^\prime = BiLSTM(C), BiLSTM(Q)$,
%	$\overleftarrow{A} = u(C^\prime, Q^\prime)$,
%	$\overrightarrow{A} = v(C^\prime, Q^\prime)$,
%	$s = m(C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%	$e = n(s, C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $C$ and $Q$ are embedding matrices for the context and the question,
%$\overleftarrow{A}$ and $\overrightarrow{A}$ are bidirectional attention flows,
%$s$ and $e$ are probabilities for starting and ending indices of the answer span.
%Our system reveals the internal states of $\overleftarrow{A}$, $\overrightarrow{A}$,
%$s$ and $e$.
The attention information is encoded as a bi-directional alignment (i.e., from context to question and vice versa). 
%The key difference is how the question to context alignment is computed (see \cite{Seo2016} for more details).
Here, we refer the context to question attention as \emph{att1}; and the question to context attention as \emph{att2} (which is a vector instead of a matrix according to \citet{Seo2016}). In this demo, we apply a min max normalization for \emph{att2} after the softmax layer to better distinguish different attention values).

@@

>>>>

\subsection{Natural Language Inference}
\label{sec:NLIexample}
The NLI task predicts the entailment relationship between a premise sentence (P) and a hypothesis sentence (H).
The attention matrix captures the alignment of words between these two sentences.
Here we give an example of how a wrong prediction can be corrected by editing attention values.

@@

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pair (P:``A woman in a green jacket is drinking tea.'' H:``A woman is drinking green tea.'') is predicted to be \emph{entailment}, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two \emph{greens} modify different nouns, which potentially leads to the wrong prediction. The grammatical structure is visually shown in the form of the dependency tree.
%i.e., \emph{green} in \textbf{P} is attached to ``jacket''.
However, the model does not have access to the syntactic information and mistakenly assumes the two \emph{greens} modify the same thing.
% (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "greens" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

\subsection{Machine Comprehension}
\label{sec:MCexample}
In the machine comprehension task, the goal is to select a span of text as the answer
to a question sentence.
%two sequences of texts are given: context and question.
%The goal is to select a span of text from the context that answers the question. 
%We provide a simple model formulation and refer the reader to the paper for details.
%%\begin{align}
%	$C^\prime, Q^\prime = BiLSTM(C), BiLSTM(Q)$,
%	$\overleftarrow{A} = u(C^\prime, Q^\prime)$,
%	$\overrightarrow{A} = v(C^\prime, Q^\prime)$,
%	$s = m(C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%	$e = n(s, C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $C$ and $Q$ are embedding matrices for the context and the question,
%$\overleftarrow{A}$ and $\overrightarrow{A}$ are bidirectional attention flows,
%$s$ and $e$ are probabilities for starting and ending indices of the answer span.
%Our system reveals the internal states of $\overleftarrow{A}$, $\overrightarrow{A}$,
%$s$ and $e$.
The attention information is encoded as a bi-directional alignment (i.e., from context to question and vice versa). 
%The key difference is how the question to context alignment is computed (see \cite{Seo2016} for more details).
Here, we refer the context to question attention as \emph{att1}; and the question to context attention as \emph{att2} (which is a vector instead of a matrix according to \citet{Seo2016}). In this demo, we apply a min max normalization for \emph{att2} after the softmax layer to better distinguish different attention values.

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "NLPVis-demo-paper"
%%% End:

### 2018-06-01 EMNLP-demo-paper/system.tex
minor edits all over

\section{Visualization System}
As illustrated in Figure~\ref{fig:modelPipeline}, many recent end-to-end NLP model follows a similar encoder, attention, and classifier structure, in which the attention stage provides a window to peek into the model decision making process. 
However, the static attention alone likely does not tell the whole story. 
The proposed system utilizes a perturbation-driven exploration strategy that allows the user manually or automatically perturb part of the pipeline and observe the changes in others.
%
The system consists of several visualization components that can be selectively combined to form a functional interactive system for a given task. In the following section, we will cover some of the essential components (the usage of all components can be found in the accompanying video \href{https://www.youtube.com/watch?v=dcpTyWfBhp0}{link}).
%such as a prediction summary statistics the entire development set, as well as a scheme to update the model based

@@

\subsection{Input Sentences Perturbation}
\label{sec:perturb}
Due to the discrete nature of the natural language, automatically applying perturbation to sentence for sensitivity analysis can be particularly challenging (compared to other domain, i.e., image), as small changes in words can leading to drastic differences in the semantics of the sentences.
To reduce the potential semantic deviation, we first devise a straightforward perturbation method by replacing \emph{nouns} and \emph{verbs} by their synonyms in the WordNet~\cite{Miller1995}. However, synonyms replacement does not guarantee the meaning of the sentence remains the same. And the WordNet often produces rare words or obscure usages that may lead to less meaningful sentences.
%However, this method requires minimal computation and is easy to implement.

To improve the perturbation quality, we adopted a translation based paraphrasing technique similar to \citet{mallinson2017paraphrasing}. Here, we translate the original English sentence into several other languages and then translate back into English. Provided the translation produces good result (in our case, we utilize the Google cloud translation platform), we can obtain paraphrases of the original sentence (see Figure~\ref{fig:sentence}, where the drop-down menu shows some of the perturbed sentences).

@@

\subsection{Prediction Visualization}
Efficient visual encoding for prediction is crucial for communicating the model behavior. And to fully support the input perturbation feature, the visual encoding should also allow multiple predictions to be shown in the same visualization.
%
For the natural language inference task, as illustrated in Figure~\ref{fig:prediction}, the predicted probabilities are encoded in the triangle barycentric coordinate, where the vertices represent the three possible predictions (namely, \emph{entailment}, \emph{contradiction}, and \emph{neutral}). The prediction for the original unperturbed input is illustrated by a larger yellow circle, whereas the prediction of perturbed inputs is represented by smaller grey circles.
A density contour of the prediction is computed to emphasize the highly cluttered areas and detect outliers.

@@

\subsection{Attention Visualization}
As illustrated in Figure~\ref{fig:attentionVis}(a)(b), the most widely adopted visual encodings for attention matrix are bipartite graph (a) and heatmap (b). 
%
In the graph attention view (Fig.~\ref{fig:attentionVis}(a)), the edge thickness corresponds to the attention value of word pairs. %The color of the rectangle text block encodes the sum of all edge values connected to it (darker shade of blues correspond to higher values).
The graph view is suitable for highlighting the most dominant alignments. However, the edges may become cluttered if multiple attention values are high. 
The matrix attention view (Fig.~\ref{fig:attentionVis}(b)) resolves these issues, despite being more verbose and less efficient in highlighting the most dominant alignments. 
We also enable the linkage between highlighted actions in both views (see Fig.~\ref{fig:attentionVis}(a)(b), where one alignment relationship is highlighted).

>>>>

\section{Visualization System}
As illustrated in Figure~\ref{fig:modelPipeline}, many recent end-to-end NLP models follow a similar encoder--attention--classifier structure. The attention stage provides a window to peek into the model decision making process. 
However, the static attention alone does not tell the whole story. 
Our proposed system uses a perturbation-driven exploration strategy that allows the user manually or automatically perturb part of the pipeline and observe the changes downstream.
%
The system consists of several visualization components that can be selectively combined to form a functional interactive system for a given task. In the following section, we will cover some of the essential components; the usage of all components can be found in the accompanying video \href{https://www.youtube.com/watch?v=dcpTyWfBhp0}{link}.
%such as a prediction summary statistics the entire development set, as well as a scheme to update the model based

@@

\subsection{Input Sentences Perturbation}
\label{sec:perturb}
Due to the discrete nature of natural language, automatically perturbing a sentence for sensitivity analysis can be particularly challenging (compared to other domains such as images) --- small changes in words can lead to drastic differences in the semantics of the sentences.
To reduce the potential semantic deviation, we allow for a straightforward perturbation method by replacing \emph{nouns} and \emph{verbs} by their synonyms in WordNet~\cite{Miller1995}. However, synonym replacement does not guarantee that the meaning of the sentence remains the same. Furthermore, WordNet often produces rare words or obscure usages that may lead to less meaningful sentences.
%However, this method requires minimal computation and is easy to implement.

To improve the perturbation quality, we also allow for a translation-based paraphrasing technique similar to \citet{mallinson2017paraphrasing}. Here, we translate the original English sentence into several other languages and then pivot back to English. Provided the translation produces good result (in our case, we use the Google cloud translation platform), we can obtain paraphrases of the original sentence (see Figure~\ref{fig:sentence}, where the drop-down menu shows some of the perturbed sentences).

@@

\subsection{Prediction Visualization}
Efficient visual encoding for prediction is crucial for communicating the model behavior. And to fully support the input perturbation feature, the visual encoding should also allow multiple predictions to be shown in the same visualization.
%
For the natural language inference task, as illustrated in Figure~\ref{fig:prediction}, the predicted probabilities are encoded via a triangular barycentric coordinate system, where the vertices represent the three possible predictions (namely, \emph{entailment}, \emph{contradiction}, and \emph{neutral}). The prediction for the original unperturbed input is illustrated by a larger yellow circle, whereas the prediction of perturbed inputs is represented by smaller grey circles.
A density contour of the prediction is computed to emphasize the highly cluttered areas and detect outliers.

@@

\subsection{Attention Visualization}
As illustrated in Figures~\ref{fig:attentionVis}(a)(b), the most widely adopted visual encodings for attention are  bipartite graphs (a) and heatmaps (b). 
%
In the graph attention view (Fig.~\ref{fig:attentionVis}(a)), the edge thickness corresponds to the attention value of the word pairs. %The color of the rectangle text block encodes the sum of all edge values connected to it (darker shade of blues correspond to higher values).
The graph view is suitable for highlighting the most dominant alignments. However, the edges may become cluttered if multiple attention values are high. 
The matrix attention view (Fig.~\ref{fig:attentionVis}(b)) resolves these issues, despite being more verbose and less efficient in highlighting the most dominant alignments. 
We also enable the linkage between highlighted actions in both views (see Fig.~\ref{fig:attentionVis}(a)(b), where one alignment relationship is highlighted).
